\documentclass[11pt]{article}
\pdfoutput=1
\usepackage[utf8]{inputenc}

% misc
\usepackage{amsmath}

% chemistry!
\usepackage[version=4]{mhchem}

% refrerences, url
\usepackage[colorlinks=true,allcolors=blue]{hyperref}

% document layout
\usepackage{indentfirst}
\setlength{\parindent}{0.4cm}
\usepackage{microtype}
\usepackage[hmarginratio=1:1,top=24mm,bottom=20mm,left=17mm,right=17mm,columnsep=20pt]{geometry}
\usepackage[pdftex]{graphicx}
\usepackage{caption}

% bibliography
\usepackage{natbib}
\bibliographystyle{plainnat}

% inline code style
\newcommand{\prog}[1]{{\tt#1}}

% TODO - do I need those?
%\usepackage{abstract} 
\usepackage{dblfloatfix} 
\usepackage{type1cm}
%\usepackage{vmargin}
\usepackage{latexsym}
\usepackage{setspace} 
\usepackage{placeins}
%\usepackage{nicefrac}
\usepackage{fancyvrb}
\usepackage{relsize}
\usepackage{lineno}

% listings
\fvset{frame=single,framerule=.2mm,fontsize=\relscale{.9}}
\input{pygments}
\newcommand*\FancyVerbStartString{}
\newcommand*\FancyVerbStopString{}

\newcounter{lstnopyt}
\setcounter{lstnopyt}{0}
\newcounter{linenopyt}
\setcounter{linenopyt}{0}

\newcounter{lstnogdl}
\setcounter{lstnogdl}{0}
\newcounter{linenogdl}
\setcounter{linenogdl}{0}

\newcommand{\codepyt}[3]{%
  \addtocounter{lstnopyt}{1}%
  \setcounter{linenopyt}{\value{FancyVerbLine}}%
  \renewcommand*\FancyVerbStartString{\PY{c}{\PYZsh{}#2}}
  \renewcommand*\FancyVerbStopString{\PY{c}{\PYZsh{}#3}}
  \setcounter{FancyVerbLine}{\thelinenopyt}%
  \fvset{label={list.P\thelstnopyt~(Python)}, rulecolor=\color{blue}}%
  \input{#1}%
}

\newcommand{\codegdl}[3]{%
  \addtocounter{lstnogdl}{1}%
  \setcounter{linenogdl}{\value{FancyVerbLine}}%
  \renewcommand*\FancyVerbStartString{\PY{c+cSingleline}{;#2}}
  \renewcommand*\FancyVerbStopString{\PY{c+cSingleline}{;#3}}
  \setcounter{FancyVerbLine}{\thelinenogdl}%
  \fvset{label={list.GDL\thelstnogdl~(Gdl/Idl)}, rulecolor=\color{black}}%
  \input{#1}%
}


% title page content
\usepackage{authblk}
\author{Anna Zimniak}
\author{Anna Jaruga}
\affil{Institute of Geophysics, Faculty of Physics, University of Warsaw, Poland}
\title{Parcel model documentation}

\begin{document}

\maketitle
\vspace{-4em}

%\twocolumn[
%  \begin{@twocolumnfalse}
%    \maketitle
%    \begin{abstract}
%    \end{abstract}
%    \vspace{2em}
%  \end{@twocolumnfalse}
%]

\section{Introduction}\label{sec:intro}

The parcel model represents an idealized scenario of a 0-dimensional air parcel 
  rising adiabatically with a~constant vertical velocity. 
Because of its simplicity, the parcel approach provides 
  computationally efficient testbed for cloud microphysics schemes.\\

Representation of microphysical and chemical processes in the parcel model 
  is done using the particle-based scheme.
The particle-based scheme (aka Lagrangian scheme) allows to track the properties of both 
  aerosol particles and cloud droplets throughout the entire simulation.
The processes resolved by the particle-based scheme cover:
  (i) condensational growth of aerosols and cloud droplets,
  (ii) collisions,
  (iii)~sedimentation\footnote{Should be switched off when used in 0-dimensional setting},
  (iv) aqueous phase chemical reactions.
All the processes can be easily switched on/off by the user.

We are using the particle-based scheme from the \emph{libcloudph++} library. 
This is a library of algorithms for representing cloud microphysics in numerical models.
The introduction to the \emph{libcloudph++} containing description 
  of the particle-based scheme used in the parcel model, 
  complete description of the programming interface of the library and a performance analysis
  is available online at \cite{Arabas_et_al_2015}.
The aqueous phase chemistry module of \emph{libcloudph++} is focused on the
  oxidation of \ce{SO2} dissolved inside water drops to sulfate.
Two reaction paths are included - oxidation by \ce{O3} and \ce{H2O2}.
A description of the representation of aqueous phase chemistry in \emph{libcloudph++} is 
  available in the PhD thesis of Jaruga.

\section{Solved equations}\label{sec:eqs}

The particle-based microphysics scheme used within the parcel model 
  expects dry air density $\rho_d$ as one of the input arguments.
To evaluate $\rho_d$, the profile of pressure within the parcel model 
  is predicted by integrating the hydrostatic equation:
\begin{equation}
	\frac{dp}{dz} = -\rho g,
	\label{hydro}
\end{equation}
\noindent
  where $p$ represents pressure, 
        $z$  is the vertical displacement, 
        $\rho$ is the density of air and
        $g$ is the gravitational acceleration.
To integrate Eq.~(\ref{hydro}) it is assumed that $\rho$ is constant
  at each height level (i.e. piecewise constant profile). 
As a result, at a given time level $n$ the pressure used to predict $\rho_d$ is defined as
  \begin{equation*}
    p^n = p^{n-1} - \rho^{n-1} g z^n .
  \end{equation*}
\noindent
The following function is used to retrieve dry air density from the pressure:
\begin{equation}
  \rho_{d}(p, \theta, r_v) = \cfrac{p - p_v(p, r_v)}{ R_d \theta (\frac{p}{p_{1000}})^{\frac{R_d}{c_{pd}}}},
  \label{rho}
\end{equation}
\noindent
where 
  $p_v(p, r_v)$ represents partial pressure of water vapor, 
  $p_{1000}$ stands for pressure equal 1000 $hPa$ that comes from the definition of potential temperature,
  $R_d$ is the gas constant for dry air and
  $c_{pd}$ is the specific heat at constant pressure for dry air.
The dry air density at a given time level $n$ is calculated as $\rho_d(p^n, \theta^n, r_v^n)$. 
The pressure outputted by the parcel model is the same 
  as the one used to calculate dry air density for the microphysics scheme.

The parcel model uses water vapor mixing ratio and dry air potential temperature
  as model variables, see \citet{Arabas_et_al_2015} for definition.
Both variables can only be changed due to the microphysical processes
  (the model is adiabatic and includes only microphysics and aqueous phase chemistry).
The derivation of source terms of heat and moisture due to microphysical processes 
  is available in Appendix A of \citep{Arabas_et_al_2015}.
For simulations with aqueous phase chemistry six additional model variables are needed
  that specify the mixing ratio of \ce{SO2}, \ce{O3}, \ce{H2O2}, \ce{CO2}, \ce{NH3} and \ce{HNO3} 
  trace gases in the model.
The aqueous phase chemistry module of \emph{libcloudph++} changes the
  trace gas mixing ratios as described in Chapter 3.3 of the PhD thesis of Jaruga.

\section{Initial condition}

The user needs to specify the initial temperature $T_0$, pressure $p_0$ and water vapor mixing ratio $r_0$.
Instead of explicitly specifying $r_0$, the user can specify the initial relative humidity $RH_0$.
The initial thermodynamic condition must be set below supersaturation, i.e.
\begin{equation}
	r_0 < \epsilon \frac{e_s(T_0)}{p_0 - e_s(T_0)},
\end{equation}
\noindent
where 
  $e_s(T_0)$ is water vapor saturation pressure at initial temperature 
  and $\epsilon = 0.622$.
The initial size distribution of dry aerosol can be a combination of any number of lognormal distributions:
\begin{equation}
        n(r) = \frac{n_{tot}}{r\sqrt{2\pi}ln(\sigma_g)}exp\left(-\frac{(ln(r)-ln(\overline{r}))^2}{2ln^2(\sigma_g)}\right)
        \label{lognormalny}
\end{equation}
\noindent
where 
  $n(r)$ is spectral density function for particles radius $r$, 
  $n_{tot}$ is total aerosol concentration, 
  $\overline{r}$ is mode radius and 
  $\sigma_g$ is geometric standard deviation.
The parameters of the size distribution can be specified by the user.
Each of the lognormal modes can have a different value of the hygroscopicity parameter $\kappa$.

For simulations with aqueous phase chemistry the initial trace gas mixing ratios 
  need to be additionally specified.
If chemical reactions are enabled, the model assumes that 
  the initial aerosol is ammonium bisulfate aerosol (\ce{NH4HSO4}).
The initial mass of chemical compounds is then calculated using the initial 
  dry aerosol size distribution and assumed dry particle density of 1.8 $g/cm^3$.
The dry particle density can be changed by the user.

\section{Arguments}

The arguments of the parcel model are divided into 5 groups:

\subsection{thermodynamic variables}

\begin{itemize}

  \item \textbf{T\_0}  : float (default = 300);\\ initial temperature [K]
  \item \textbf{p\_0}  : float (default = 101300);\\ initial pressure [Pa]
  \item \textbf{r\_0}  : float (default = 0.022);\\ initial water vapor mass mixing ratio [kg/kg].
  \item \textbf{RH\_0} : float (no default value);\\ initial relative humidity, alternative to the \textbf{r\_0} option.
  \item \textbf{w} : float (default = 1);\\ Updraft velocity [m/s]

\end{itemize}

\subsection{chemistry module variables}

\begin{itemize}

  \item \textbf{SO2\_g}    : float (default = 0); \\ initial \ce{SO2} trace gas mixing ratio [kg/kg]
  \item \textbf{O3\_g}     : float (default = 0); \\ initial \ce{O3} trace gas mixing ratio [kg/kg]
  \item \textbf{H2O2\_g}   : float (default = 0); \\ initial \ce{H2O2} trace gas mixing ratio [kg/kg]
  \item \textbf{CO2\_g}    : float (default = 0); \\ initial \ce{CO2} trace gas mixing ratio [kg/kg]
  \item \textbf{NH3\_g}    : float (default = 0); \\ initial \ce{NH3} trace gas mixing ratio [kg/kg]
  \item \textbf{HNO3\_g}   : float (default = 0); \\ initial \ce{HNO3} trace gas mixing ratio [kg/kg]
  \item \textbf{chem\_rho} : float (default = 1800);  \\ assumed dry particulate density [kg/m3]
  \item \textbf{chem\_dsl} : bool  (default = false); \\ switch for dissolving trace gases into droplets
  \item \textbf{chem\_dsc} : bool  (default = false); \\ switch for dissociation of chemical compounds in droplets
  \item \textbf{chem\_rct} : bool  (default = false); \\ switch for aqueous phase reactions in droplets

\end{itemize}

\subsection{aerosol attributes}

\begin{itemize}
  \item \textbf{aerosol} : jason string (default =\\ '\{"ammonium\_sulfate": \{"kappa": 0.61, "mean\_r": [0.02e-6], "gstdev": [1.4], "n\_tot": [60.0e6]\}\}');\\ 

    \vspace{-.9em}
    It is a Python dictionary of dictionaries defining initial dry aerosol spectrum.
    First key defines the name of the aerosol type. The following dictionary defines the spectrum of aerosol of this type:\\
      \indent $\cdot$\ "kappa" - hygroscopicity parameter of the aerosol (see \citep{Petters_et_al_2007} for more details)\\
      \indent $\cdot$\ "mean\_r" - list of mean radii of lognormal modes of aerosol of this type [m]\\
      \indent $\cdot$\ "gstdev" - list of geometric standard deviations of lognormal modes of aerosol of this type \\
      \indent $\cdot$\ "n\_tot" - list of concentrations (at the STP dry air density $\rho_d = 1.225$ kg$/$m$^3$) of lognormal modes of aerosol of this type [m\textsuperscript{-3}] \\

    \vspace{-.9em}
    For example, user can define two types of aerosol - bimodal ammonium bisulfate and monomodal sea salt : \\ 
  
$$
\mbox{"ammonium\_bisulfate"}: \left\{ \begin{array}{ll}
\mbox{"kappa": 0.61}  & \textrm{}\\
\mbox{"mean\_r": [0.02e-6, 0.07e-7]}  & \textrm{}\\
\mbox{"gstdev": [1.4, 1.2]} & \textrm{}\\
\mbox{"n\_tot": [120.0e6, 80.0e6]} & \textrm{}\\
\end{array} \right.
$$

$$
\mbox{"sea\_salt"}: \left\{ \begin{array}{ll}
\mbox{"kappa": 1.28}  & \textrm{}\\
\mbox{"mean\_r": [2e-6]}  & \textrm{}\\
\mbox{"gstdev": [1.6]} & \textrm{}\\
\mbox{"n\_tot": [1e3]} & \textrm{}\\
\end{array} \right.
$$

\end{itemize}

\subsection{simulation parameters}\label{sec:simpar}

\begin{itemize}

  \item \textbf{dt} : float (default = 0.1); \\ timestep [s]
  \item \textbf{z\_max} : float (default = 200); \\ maximum vertical displacement [m]
  \item \textbf{sd\_conc} : int (default = 64); \\ number of super-droplets used by the particle-based microphysics scheme
  \item \textbf{pprof} : string (default = "pprof\_piecewise\_const\_rhod"); \\ 
                method to calculate pressure profile used to calculate 
                dry air density that is used by the super-droplet scheme. The parcel model uses method described in sec. \ref{sec:eqs}. It is possible to switch on another option: "pprof\_const\_th\_rv" (see Appendix A for details).
\end{itemize}

\subsection{output parameters}\label{sec:output}

\begin{itemize}
  \item \textbf{outfile} : string (default = "test.nc"); \\ output file name; the output file is in NetCDF format
  \item \textbf{outfreq} : int (default = 100); \\ output frequency (time gap between outputted points in number of time steps)
  \item \textbf{out\_bin} : jason string (default =\\ '\{"radii": \{"rght": 0.0001, "moms": [0], "drwt": "wet", "nbin": 26, "lnli": "log", "left": 1e-09\}\}');\\

    \vspace{-.9em}
    It is a Python dictionary of dictionaries defining spectrum diagnostics.
    First key defines the name of created variable. The following dictionary defines spectrum characteristics:\\
      \indent $\cdot$\ "rght", "left" - right and left edge of the spectrum (in meters)\\
      \indent $\cdot$\ "moms" - list of numbers, specifying moments of the spectrum\\
      \indent $\cdot$\ "drwt" - choice between  ("dry") for dry aerosol spectrum and ("wet") for wet particles\\
      \indent $\cdot$\ "nbin" - number of bins\\
      \indent $\cdot$\ "lnli" - linear ("lin") or logaritmical ("log") spacing between between bins\\
 
    \vspace{-.9em}
    For example, user can define two variables - "A" and "B" : \\ 
  
$$
\mbox{"A"}: \left\{ \begin{array}{ll}
\mbox{"rght": 0.0001}  & \textrm{}\\
\mbox{"moms": [0]}  & \textrm{}\\
\mbox{"drwt": "wet"} & \textrm{}\\
\mbox{"nbin": 26} & \textrm{}\\
\mbox{"lnli": "log"}  & \textrm{}\\
\mbox{"left": 1e-09} & \textrm{}\\
\end{array} \right.
$$

$$
\mbox{"B"}: \left\{ \begin{array}{ll}
\mbox{"rght": 2.5e-05} & \textrm{}\\
\mbox{"moms": [0, 1, 2, 3] } & \textrm{}\\
\mbox{"drwt": "wet"} & \textrm{}\\
\mbox{"nbin": 49} & \textrm{}\\
\mbox{"lnli": "lin"} & \textrm{}\\
\mbox{"left": 5e-07} & \textrm{}\\
\end{array} \right.
$$
 
\vspace{0.35cm}
   It will generate five output spectra:
      0-th spectrum moment for 26 bins spaced logaritmically 
        between $10^{-9}$ and $10^{-4}$ m for wet radius for variable 'A' and 
      0, 1, 2 and 3-rd moments for 49 bins spaced linearly 
        between $5\cdot10^{-7}$ and $2.5\cdot10^{-5}$ m for wet radius for variable 'B'.

   \vspace{1.5em}
   For simulations with aqueous phase chemistry the ``moms'' key can also be used 
     for outputting the total mass of chemical compounds dissolved in droplets.
   The allowed values for moms are "SO2\_a", "CO2\_a", "NH3\_a", "HNO3\_a", "O3\_a", "H2O2\_a", "S\_VI" and "H".
   The first six output the total dissolved mass of the corresponding trace gases.
   The "S\_VI" outputs the mass of created \ce{H2SO4} and the "H" outputs the mass of \ce{H^+} ions.
   The number of bins and the size range of droplets selected for output 
     is defined in the same way as for the previously described output of size distribution moments.

\end{itemize}

\section{Output}

The parcel model uses NetCDF file format\footnote{see unidata.ucar.edu/software/netcdf/ for details} for output.
The content of the output file can be be viewed in terminal by using the \prog{ncdump} command.
All the initial parameters of the parcel model are saved as global attributes of the output file.
Additionally the maximum relative humidity reached during simulation 
  \textbf{RH\_{max}} is also saved as a global attribute.
Output variables describing time-dependent ambient conditions in the parcel are saved as:
\begin{itemize}
  \item \textbf{t} - time [s],
  \item \textbf{z} - height above the starting point[m],
  \item \textbf{p} - pressure [Pa],
  \item \textbf{r\_v} - water vapor mixing ratio [kg/kg],
  \item \textbf{RH} - relative humidity [1],
  \item \textbf{T} - temperature [K],
  \item \textbf{th\_d} - dry potential temperature [K],
  \item \textbf{SO2\_g, O3\_g, H2O2\_g, CO2\_g, NH3\_g, HNO3\_g} - trace gas mixing ratios [kg/kg of dry air]
  \item \textbf{SO2\_a, O3\_a, H2O2\_a, CO2\_a, NH3\_a, HNO3\_a} - total number of moles dissolved in droplets [moles/kg of dry air]
\end{itemize}
\noindent
Output variables describing size distribution of particles or mass of the dissolved chemical compounds 
  from the particle tracking scheme depend on the user-defined "out\_bin" parameter.
For example, for the default setting of "out\_bin" parameter there are two output variables 
  describing placing and spacing of size distribution bins:
\begin{itemize}
  \item \textbf{radii\_r\_wet} - left edges of the bins of the spectrum histogram [m],
  \item \textbf{radii\_dr\_wet} - bins width [m]
\end{itemize}
and one output variable containing the time-dependent chosen moment of size distribution:
\begin{itemize}
  \item \textbf{radii\_m0} - 0th moment of spectrum.
\end{itemize}

\section{Installation}

The parcel model requires the \emph{libcloudph++} library to be installed. 
The \emph{libcloudph++} library can be obtained from the project Github repository 
  at \url{https://github.com/igfuw/libcloudphxx}.
The library dependencies and installation guidelines are available there in a Readme file.

The parcel model is available online at the project repository at \url{https://github.com/igfuw/parcel}.
The installation guidelines are available there in a Readme file.
The parcel model is written in Python 2.7.

\section{Usage examples}

\subsection{bash}

\noindent
To run the model using default parameters type in terminal:

\vspace{0.35cm}
  \prog{ python parcel.py}
\vspace{0.35cm}

\noindent
Any arguments to parcel model options can be passed via command line. For example:

\vspace{0.35cm}
 \prog{python parcel.py \texttt{--}outfile 'test2.nc' \texttt{--}p\_0 100000 \texttt{--}T\_0 280 \texttt{--}r\_v0 0.025 \texttt{--}out\_bin '\{"radii": \{"rght": 0.0001,  "moms": [0], "drwt": "wet", "nbin": 26, "lnli": "log", "left": 1e-09\}, "cloud": \{"rght": 2.5e-05, "moms": [0, 1, 2, 3], "drwt": "wet", "nbin": 49, "lnli": "lin", "left": 5e-07\}\}'}
\vspace{0.35cm}

\noindent
will generate an output file 'test2.nc', which contains results of the simulation with initial pressure 1000~hPa, initial temperature 280 K, initial water vapor mixing ratio 25 g/kg and two spectrum variables - 'radii' and 'cloud' (the same as described in sec. \ref{sec:output}). Other arguments will be default.

\subsection{py.test}

The parcel model is shipped with a set of tests designed for checking the particle-based microphysics 
  and aqueous phase chemistry schemes from the \emph{libcloudph++} library.
The tests may serve as usage examples for the parcel model and \emph{libcloudph++} library.
The tests are located in the {\bf unit\_test} and {\bf long\_test} folders.
The tests use the py.test Python package for test automation and can be run by typing from terminal:\\

\prog{py.test unit\_test}\\

\noindent
To enable more output during testing type:\\

\prog{py.test -s -v long\_test}\\

\noindent
Example output plots generated by the tests are saved in the \prog{plots/outputs} folder.

\newpage

\subsection{Gdl/Idl}

The following example shows how to use the parcel model from the IDL (or its free counterpart GDL)
   programming languages. 
It takes advantage of the command line interface of the parcel model
  and executes the simulation via the \prog{spawn} command.
It then plots the size distribution of particles at different timelevels.

% first column
\begin{minipage}[t]{0.95\textwidth}
\codegdl{lst/plot.pro}{listing00}{listing01}
\end{minipage}
%second column
\begin{minipage}[t]{0.3\textwidth}
\end{minipage}

\newpage
\subsection{Python}

The following example shows how to use the parcel model from Python programming language 
  and plot vertical profiles for pressure, temperature and relative humidity from NetCDF file. 

% first column
\begin{minipage}[t]{0.65\textwidth}
\codepyt{lst/plot.py}{listing00}{listing01}
\end{minipage}
%second column
\begin{minipage}[t]{0.3\textwidth}
\end{minipage}

\appendix

\section*{\\Appendix A} \label{App:AppendixA}

Another method for integrating equation (\ref{hydro}) and retrieving pressure profile is to
assume constant potential temperature, $\theta = \theta^0$, and constant water vapor mixing ratio, $r_v = r_v^0$ (option "pprof\_const\_th\_rv", see sec. \ref{sec:simpar}).

  As a result at a given time level $n$ the pressure used to predict $\rho_d$ is defined as
  \begin{equation*}
    p^n = p_{1000} \left( (\frac{p^0}{p_{1000}})^{\frac{R_d}{c_{pd}}} - \frac{R_d g (z^n - z^0)}{c_{pd} \theta^{0} R(r_v^0)} \right) ^ {\frac{c_{pd}}{R_d}} ,
  \end{equation*}
  \noindent
  where $p_{1000}$ stands for pressure equal 1000 $hPa$ that comes from the definition of potential temperature,
        $R(r_v^0)$ is the gas constant for moist air and
        $c_{pd}$ is the specific heat at constant pressure for dry air.

The function (\ref{rho}) is used to retrieve dry air density from pressure. The dry air density at a given time level $n$ is calculated as $\rho_d(p^n, \theta^0, r_v^0)$.
The pressure outputted by the parcel model is then calculated as $ \rho_d  (R_d + r_v R_v)  T$,
  where $R_v$ is the gas constant for water vapor.

This method follows the procedure used in the 2-dimensional kinematic model \emph{icicle}, 
  described in \citep{Arabas_et_al_2015} and was added to the parcel model to allow better comparison 
  between the two setups.

\bibliography{doc}

\end{document}
